---
title: "Neutralization assays"
author: "Spyros"
date: "3/8/2019"
output: html_document
---

```{r setup, include=FALSE}
require(drc)
library(gridExtra)
library(grid)
library(gplots)
require(ggthemes)
```

Import data 
```{r Import}
data <- read.csv("~/Documents/Labwork/Projects/Biohub/DRC_neutralization/input/LG119_Plate2.csv", sep=",", skip = 1)
data$Well.ID <- as.character(data$Well.ID)
head(data)
```

I assume the format is cols 1-11 and cols 14-23 with 11 and 23 being the no antibody controls 
I also assume that rows are paired BC, DE, etc .. 
First row is B and last row is O
Pairs of wells are in the same row 
```{r Prepare pairs matrix}
# Crete a matrix of row pairs
rows <- LETTERS[2:15]
pairs <- matrix(rows, ncol=2, byrow = T)
pairs
```

Set the concentration vector in ng/ml
```{r Prepare  concentration vector}
conc <- c(10000,2000, 400, 80, 16, 3.2, 0.64, 0.128, 0.0256)
conc.log <- log10(conc)
conc
```

```{r Create list to store output}
# Create list to store regression models and ggplots 
list.reg <- list()
list.eds <- list()
list.gfp.max <- list()
list.rsd <- list()
list.means <- list()
list.sd <- list()
```

Provide a vector of antibody names 
For now its just AB1, AB2, etc ..
```{r Provide names for each assay}
abs.left <- paste("AB_L",1:nrow(pairs), sep="_")
abs.right <- paste("AB_R",1:nrow(pairs), sep="_")
```

Create color wheel
```{r Create color wheel}
col.wheel <- tableau_color_pal(palette = "tableau10")(10)
# Plot along concentration 
pie(rep(0.1,10), col=col.wheel)
```


# Process the left half of the plate
```{r Left half of plate}
for(i in 1:nrow(pairs)){
  # Subset data from main data frame 
  pair.1 <- data[grep(pairs[i,1], data$Well.ID),][1:10,] # 1:10 selects X02-X11
  pair.2 <- data[grep(pairs[i,2], data$Well.ID),][1:10,]
  # Get GFP value and normalize to no antibody control (last entry X11)
  # grep for 11 to get the X11 column from both replicates
  con.pair.1 <- pair.1$X.GFP.[grep(11, pair.1$Well.ID)] 
  con.pair.2 <- pair.2$X.GFP.[grep(11, pair.2$Well.ID)]
  # Correct the GFP values by dividing with the no antibody control 
  cor.pair.1.t <- pair.1$X.GFP./con.pair.1
  cor.pair.2.t <- pair.2$X.GFP./con.pair.2
  # Remove the zero value 
  cor.pair.1 <- cor.pair.1.t[-length(cor.pair.1.t)]
  cor.pair.2 <- cor.pair.2.t[-length(cor.pair.2.t)]
  # Get average values 
  cor.means <- rowMeans(cbind(cor.pair.1, cor.pair.2))
  cor.st <- apply(cbind(cor.pair.1, cor.pair.2), MARGIN = 1, range)
  # Store in list 
  list.means[[i]] <- cor.means
  list.sd[[i]] <- cor.st
  # DOES IT MAKE SENSE ? 
  if(length(which(is.infinite(cor.means)==T)) ==0){
    # Fit using a 4-parameter log logistic model and store the model
    list.reg[[i]] <- drm(cor.means ~ conc,fct = LL.4())
    # Get ED5,10 and 50 values 
    list.eds[[i]] <- ED(object = list.reg[[i]], c(5,10,50), interval = "delta", display=F)
    # Also store %GFP and max antibody 
    list.gfp.max[[i]] <- cor.means[1]*100
    # Also store the residual standard error 
    list.rsd[[i]] <- abs(sqrt(summary(list.reg[[i]])$"resVar") / mean(fitted(list.reg[[i]])))
    # Plot
    plot(list.reg[[i]], col=col.wheel[i], 
         xlab="MAb Concentration (ng/ml)",
         ylab="% Infection", ylim=c(0,1.5), 
         pch=19, cex=0.5, axes = F)
    suppressWarnings(arrows(conc, cor.st[1,], conc, cor.st[2,], length=0.05, angle=90, code=3, lty=3, col=col.wheel[i]))
    par(new=T)
  }}
legend("right", "top", legend=abs.left, fill=col.wheel[1:length(abs.left)]) 
axis(side=1, c(0.01,0.1,1,10,100,1000,10000))
axis(side=2, seq(0,1.5,0.1))
```


Make and plot table (Left)
```{r }
# Name the listss 
names(list.reg) <- abs.left
names(list.eds) <- abs.left
# Make table of assay stats
mat.plot <- as.data.frame(matrix(nrow=length(list.reg), ncol=5))
colnames(mat.plot) <- c("ab ID", "ED50", "ED50 error", "%GFP at ab max", "RSD")
for(i in 1:length(list.reg)){
  mat.plot[i,1] <- abs.left[i]
  mat.plot[i,2] <- round(list.eds[[i]][3,1],2) # get ED 50 
  mat.plot[i,3] <- round(list.eds[[i]][3,2],2) # get ED 50 st. error 
  mat.plot[i,4] <- round(list.gfp.max[[i]],2)
  mat.plot[i,5] <- round(list.rsd[[i]],3)
}
# Plot table 
tbl <- tableGrob(mat.plot)
plot(tbl)
```

Plot each assay individually (Left)
```{r }
par(mfcol=c(3,3), mar=c(5,5,3,3))
for(i in 1:length(list.reg)){
  plot(list.reg[[i]], col=col.wheel[i], 
       xlab="MAb Concentration (ng/ml)",
       ylab="% Infection", ylim=c(0,1.5), 
       pch=19, cex=0.5, axes = T, lty=3, main=names(list.reg)[i])
  suppressWarnings(arrows(conc, list.sd[[i]][1,], conc, list.sd[[i]][2,], length=0.05, angle=90, code=3, lty=1, col=col.wheel[i]))
  # Add ED50 and SDR on the plot 
  text(x=5, y=1.4, paste("ED50=",mat.plot[i,"ED50"],sep=""),cex = 0.7)
  text(x=5, y=1.3, paste("RSD=",mat.plot[i,"RSD"],sep=""),cex = 0.7)
  }
```


# Process the right half of the plate
```{r message=FALSE}
for(i in 1:nrow(pairs)){
  # Subset data from main data frame 
  pair.1 <- data[grep(pairs[i,1], data$Well.ID),][1:10,] # 1:10 selects X02-X11
  pair.2 <- data[grep(pairs[i,2], data$Well.ID),][1:10,]
  # Get GFP value and normalize to no antibody control (last entry X11)
  # grep for 11 to get the X11 column from both replicates
  con.pair.1 <- pair.1$X.GFP.[grep(11, pair.1$Well.ID)] 
  con.pair.2 <- pair.2$X.GFP.[grep(11, pair.2$Well.ID)]
  # Correct the GFP values by dividing with the no antibody control 
  cor.pair.1.t <- pair.1$X.GFP./con.pair.1
  cor.pair.2.t <- pair.2$X.GFP./con.pair.2
  # Remove the zero value 
  cor.pair.1 <- cor.pair.1.t[-length(cor.pair.1.t)]
  cor.pair.2 <- cor.pair.2.t[-length(cor.pair.2.t)]
  # Get average values 
  cor.means <- rowMeans(cbind(cor.pair.1, cor.pair.2))
  cor.st <- apply(cbind(cor.pair.1, cor.pair.2), MARGIN = 1, range)
  # Store in list 
  list.means[[i]] <- cor.means
  list.sd[[i]] <- cor.st
  # DOES IT MAKE SENSE ? 
  if(length(which(is.infinite(cor.means)==T)) ==0){
    # Fit using a 4-parameter log logistic model and store the model
    list.reg[[i]] <- drm(cor.means ~ conc,fct = LL.4())
    # Get ED5,10 and 50 values 
    list.eds[[i]] <- ED(object = list.reg[[i]], c(5,10,50), interval = "delta", display=F)
    # Also store %GFP and max antibody 
    list.gfp.max[[i]] <- cor.means[1]*100
    # Also store the residual standard error 
    list.rsd[[i]] <- abs(sqrt(summary(list.reg[[i]])$"resVar") / mean(fitted(list.reg[[i]])))
    # Plot
    plot(list.reg[[i]], col=col.wheel[i], 
         xlab="MAb Concentration (ng/ml)",
         ylab="% Infection", ylim=c(0,1.5), 
         pch=19, cex=0.5, axes = F)
    suppressWarnings(arrows(conc, cor.st[1,], conc, cor.st[2,], length=0.05, angle=90, code=3, lty=3, col=col.wheel[i]))
    par(new=T)
  }}
legend("right", "top", legend=abs.right, fill=col.wheel[1:length(abs.right)]) 
axis(side=1, c(0.01,0.1,1,10,100,1000,10000))
axis(side=2, seq(0,1.5,0.1))
```

Make and plot table (Right)
```{r }
# Name the listss 
names(list.reg) <- abs.right
names(list.eds) <- abs.right
# Make table of assay stats
mat.plot <- as.data.frame(matrix(nrow=length(list.reg), ncol=5))
colnames(mat.plot) <- c("ab ID", "ED50", "ED50 error", "%GFP at ab max", "RSD")
for(i in 1:length(list.reg)){
  mat.plot[i,1] <- abs.right[i]
  mat.plot[i,2] <- round(list.eds[[i]][3,1],2) # get ED 50 
  mat.plot[i,3] <- round(list.eds[[i]][3,2],2) # get ED 50 st. error 
  mat.plot[i,4] <- round(list.gfp.max[[i]],2)
  mat.plot[i,5] <- round(list.rsd[[i]],3)
}
# Plot table 
tbl <- tableGrob(mat.plot)
plot(tbl)
```

Plot each assay individually (Right)
```{r }
par(mfcol=c(3,3), mar=c(5,5,3,3))
for(i in 1:length(list.reg)){
  plot(list.reg[[i]], col=col.wheel[i], 
       xlab="MAb Concentration (ng/ml)",
       ylab="% Infection", ylim=c(0,1.5), 
       pch=19, cex=0.5, axes = T, lty=3, main=names(list.reg)[i])
  suppressWarnings(arrows(conc, list.sd[[i]][1,], conc, list.sd[[i]][2,], length=0.05, angle=90, code=3, lty=1, col=col.wheel[i]))
  # Add ED50 and SDR on the plot 
  text(x=5, y=1.4, paste("ED50=",mat.plot[i,"ED50"],sep=""),cex = 0.7)
  text(x=5, y=1.3, paste("RSD=",mat.plot[i,"RSD"],sep=""),cex = 0.7)
}
```


  
  
  
  
